### 2023-10-25 generalExploration.Rmd
add death handling

source("R/population.R")
N=100
z=4 #number of neutral traits
ki <- 5
km <- 2
K  <-  ki+km
m=.2 #proba marriage
b=.1 #HG birth rate
r=.2 #farmer scale param
rho #marriage rule

@@

reproduction <- function(p1,p2,tp,i=NULL){
    print(i)
    stopifnot(p1$community==p2$community)
    traits=ifelse(sample(c(0,1),length(tp$s),replace=T),p1$traits,p2$traits)
    list(
         id=i,
         traits=traits,
         age=0,
         community=p1$community,
         sex=sample(c("F","M"),1),
         repro=FALSE,
         partner=-1
    )
}

population=lapply(1:N,function(i)
                      list(
                           id=i,
                           traits=sample(c(0,1),size=z,replace=T),
                           age=sample.int(80,1),
                           community=sample.int(K,1),
                           sex=sample(c("F","M"),1),
                           repro=FALSE,
                           partner=-1
                      )
)

@@

tstep=1000
popsize=c()
popsum=list()
for(time in 1:tstep){
    for(i in sample(1:length(population))){
        if(!is.null(population[[i]])){
            ind=population[[i]]$id
            ##ageing
            population[[ind]]$age=population[[ind]]$age+1

            ##repro
            partner=population[[ind]]$partner
            repro=population[[ind]]$repro
            if(partner>0 & !repro){
                print(paste("childtest",ind,partner,population[[ind]]$community,population[[partner]]$community))
                population[[partner]]$repro=TRUE #make sure proba of reproduction is unique for all pair 
                ad_tr=initcomus$adaptivetraits[population[[ind]]$community,]
                lambda=b+r*sum(ad_tr)
                nchild=rpois(1,lambda)
                if(nchild>0){
                    for(i in 1:nchild){
                        print("newborn")
                        population[[length(population)+1]]=reproduction(population[[ind]],population[[partner]],neutraltraitsParam,length(population)+1)
                    }
                }
            }
            population[[ind]]$repro=FALSE

            ##marriage
            if(population[[ind]]$age>18 & population[[ind]]$partner <0){
                print("hmmmaybe")
                if(runif(1)<m){
                    potential=sapply(population,function(i,s)if(i$age>18 & i$partner<0 & i$sex != s )return(i$id),s=population[[ind]]$sex)
                    potential=unlist(potential)
                    if(length(potential)>0){
                        if(length(potential)==1)partner=potential
                        else partner=sample(potential,1)
                        population[[ind]]$partner=partner
                        population[[partner]]$partner=ind
                        newcom=ifelse(runif(1)<rho,population[[ind]]$community,population[[partner]]$community)
                        population[[ind]]$community=population[[partner]]$community=newcom
                        print(paste("marriage",ind,partner,"moving all to",newcom,":",population[[partner]]$community,population[[ind]]$community))
                    }
                }
            }
            #if(runif(1)<d)population[[ind]]=NULL
        }

@@

        ##pm transmission
        #quickly summarize
        popsum[[time]]=sapply(names(population[[1]]),function(n)table(sapply(population,"[[",n)))
        size=table(sapply(population,"[[","community"))
        initcomus$size=size
        ##
        plot(initcomus$coordinates,pch=21,bg=apply(initcomus$adaptivetraits,1,mean)+1,cex=log(initcomus$size))
        popsize=c(popsize,length(population))
    }
}

>>>>

source("R/population.R")
N=100
z=4 #number of neutral traits
ki <- 5
km <- 2
K  <-  ki+km
m=.2 #proba marriage
b=.1 #HG birth rate
r=.2 #farmer scale param
rho=.5 #marriage rule
d=0.01

@@

reproduction <- function(p1,p2,tp,i=NULL){
    stopifnot(p1$community==p2$community)
    traits=ifelse(sample(c(0,1),length(tp$s),replace=T),p1$traits,p2$traits)
    if(!is.null(i))id=as.numeric(i)+1
    print(id)
    list(
         id=as.character(id),
         traits=traits,
         age=0,
         community=p1$community,
         sex=sample(c("F","M"),1),
         repro=FALSE,
         partner=-1
    )
}

ip=1:N
names(ip)=ip
population=lapply(ip,function(i)
                      list(
                           id=as.character(i),
                           traits=sample(c(0,1),size=z,replace=T),
                           age=sample.int(80,1),
                           community=sample.int(K,1),
                           sex=sample(c("F","M"),1),
                           repro=FALSE,
                           partner=-1
                      )
)

@@

tstep=1000
popsize=c()
popsum=list()
for(time in 1:tstep){
    for(i in sample(1:length(population))){
        ind=tryCatch(population[[i]]$id,error=function(e)NULL)
        if(!is.null(ind)){
            ##ageing
            population[[ind]]$age=population[[ind]]$age+1

            ##repro
           partner=population[[ind]]$partner
            repro=population[[ind]]$repro
            if(partner>0 & !repro){
                print(paste("childtest",ind,partner,population[[ind]]$community,population[[partner]]$community))
                population[[partner]]$repro=TRUE #make sure proba of reproduction is unique for all pair 
                ad_tr=initcomus$adaptivetraits[population[[ind]]$community,]
                lambda=b+r*sum(ad_tr)
                nchild=rpois(1,lambda)
                if(nchild>0){
                    for(i in 1:nchild){
                        print("newborn")
                        newborn=reproduction(population[[ind]],population[[partner]],neutraltraitsParam,population[[length(population)]]$id)
                        population[[newborn$id]]=newborn
                    }
                }
            }
            population[[ind]]$repro=FALSE

            ##marriage
            if(population[[ind]]$age>18 & population[[ind]]$partner <0){
                print("marriage attempt")
                if(runif(1)<m){
                    potential=sapply(population,function(i,s)if(i$age>18 & i$partner<0 & i$sex != s )return(i$id),s=population[[ind]]$sex)
                    potential=unlist(potential)
                    print(potential)
                    if(length(potential)>0){
                        if(length(potential)==1)partner=potential
                        else partner=sample(potential,1)
                        population[[ind]]$partner=partner
                        population[[partner]]$partner=ind
                        newcom=ifelse(runif(1)<rho,population[[ind]]$community,population[[partner]]$community)
                        population[[ind]]$community=population[[partner]]$community=newcom
                        print(paste("marriage",ind,partner,"moving all to",newcom,":",population[[partner]]$community,population[[ind]]$community))
                    }
                }
            }
            if(runif(1)<d){
                if(partner>0) population[[partner]]$partner=-1
                population[i]=NULL
                print(paste("dead ind",i))
            }
        }

@@

        ##pm transmission
        #quickly summarize
        popsum[[time]]=sapply(names(population[lengths(population)>1][1]),function(n)table(sapply(population,"[[",n)))
        size=table(sapply(population,"[[","community"))
        initcomus$size=size
        ##
        plot(initcomus$coordinates,pch=21,bg=apply(initcomus$adaptivetraits,1,mean)+1,cex=log(initcomus$size))
        popsize=c(popsize,length(population))
    }
}
